---
title: "rISIMIP - R package for accessing and analysing ISIMIP Environmental Data"
output: 
  github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, warning=F, comment=NA, message=F, fig.path="figures/")
```

## Overview

`rISIMIP` is an R package for <!--downloading,--> accessing and analysing data provided by the Inter-sectoral Impact Model Intercomparison Project (ISIMIP). Data from the different simulation rounds (ISIMIP2a, ISIMIP2b, ISIMIP3a, ISIMIP3b) is available from [here](https://esg.pik-potsdam.de/search/isimip/). For more information on the different data types and data input and output products have a look at the [ISIMIP Website](https://www.isimip.org/). The package currently consists of two functions:

<!--* `getISIMIP()` downloads ISIMIP data -->
<!-- getISIMIP should call listISIMIP() and than run download!!!-->
* `readISIMIP()` reads and pre-processes ISIMIP data
* `listISIMIP()` creates a list of requested ISIMIP data files

You can learn more about them in `vignette("rISIMIP")`. 

An example of extracting country-specific data from ISIMIP2b can be found in `vignette("country-specific")`.

## Installation

To *use* the package, install it directly from GitHub using the `remotes` package:

```{r}
# Install remotes if not previously installed
if(!"remotes" %in% installed.packages()[,"Package"]) install.packages("remotes")

# Install rISIMIP from Github if not previously installed
if(!"rISIMIP" %in% installed.packages()[,"Package"]) remotes::install_github("RS-eco/rISIMIP", build_vignettes = TRUE)
```

**If you encounter a bug or if you have any problems, please file an [issue](https://github.com/RS-eco/rISIMIP/issues) on Github.**

## Usage

```{r}
# Load rISIMIP package
library(rISIMIP)
```

### List ISIMIP files

The function `listISIMIP` just lists all climate files for the desired time period, model and variable. The files can then be put into the `aggregateNC` function of the `processNC` package for processing the required NetCDF files.

```{r list_urbanareas, eval=F}
# List urban area files for histsoc scenario - ISIMIP2b
listISIMIP(path="I:/", version="ISIMIP2b", type="landuse", 
           scenario="histsoc", var="urbanareas", startyear=1861, endyear=2005)

# List crop data files for histsoc scenario - ISIMIP3b
listISIMIP(path="I:/", version="ISIMIP3b", type="landuse", 
           scenario="histsoc", var="5crops", startyear=1861, endyear=2005)
```

**Note:** The path must lead to a file directory on your computer, which contains the required ISIMIP files. You can download the required ISIMIP data files from: https://esg.pik-potsdam.de/search/isimip/

### Read ISIMIP files

With `readISIMIP` you can read one or multiple ISIMIP datafiles into a raster stack.

```{r read_urbanareas, eval=F}
# Read urban area file for 2005soc scenario - ISIMIP2b
(urbanareas_1970_1999 <- readISIMIP(path="I:/", type="landuse", scenario="2005soc", 
                                    var="urbanareas", startyear=1970, endyear=1999))

# Read pasture file for 2015soc scenario - ISIMIP3b
(pastures_1970_1999 <- readISIMIP(path="I:/", type="landuse", scenario="2015soc", 
                                    var="pastures", startyear=1970, endyear=1999))
```

However, this is not useful if you are interested in long time periods, as one datafile is about 7 GB in size and you will quickly run into memory limitations.

## Internal data

`rISIMIP` contains various pre-processed data. See the *data-raw* folder for how we derived the included datasets.

### Temperature thresholds

Annual global mean temperature as well as the 31-year running mean were calculated for each GCM and four RCPs (RCP2.6, RCP4.5, RCP6.0 and RCP8.5). Furthermore, the year when the 31-year runnning mean of global mean temperature crosses a certain temperature threshold has been calculated. The data has been provided by ISIMIP and a summary of it can be accessed from the `vignette("temperature-thresholds")` vignette and is also available from [the ISIMIP Website](https://www.isimip.org/protocol/temperature-thresholds-and-time-slices/).

### Landseamask

The landseamask used by ISIMIP has been included in the package and can be accessed by:

```{r}
data("landseamask_generic")
```

### Bioclimatic data

The code for calculating global bioclimatic data from ISIMIP2b and ISIMIP3b mnodel output can be found in `vignette("global-landonly")` and `vignette("global-landonly-isimip3b")` respectively.

#### ISIMIP2b & EWEMBI

Current and future bioclimatic data for three 30-yr periods (1995, 2050, 2080) was derived from the EWEMBI (https://esg.pik-potsdam.de/search/isimip/?project=ISIMIP2b&amp;product=input_secondary&amp;dataset_type=Climate+atmosphere+observed) and ISIMIP2b data (https://esg.pik-potsdam.de/search/isimip/?project=ISIMIP2b&amp;product=input&amp;dataset_type=Climate+atmosphere+simulated) and is included in this package. 

**EWEMBI - 1995**

```{r bio1_1995, fig.width=9, fig.height=4}
data("bioclim_ewembi_1995_landonly")

library(dplyr); library(sf); library(ggplot2)
data(outline, package="ggmap2")
outline <- sf::st_as_sf(outline)
col_val <- scales::rescale(unique(c(seq(min(bioclim_ewembi_1995_landonly$bio1), 0, length=5),
                                    seq(0, max(bioclim_ewembi_1995_landonly$bio1), length=5))))

bioclim_ewembi_1995_landonly %>% select(x,y,bio1) %>% 
  ggplot() + geom_tile(aes(x=x, y=y, fill=bio1)) + 
  geom_sf(data=outline, fill="transparent", colour="black") + 
  scale_fill_gradientn(name="tmean (°C)", colours=rev(colorRampPalette(
    c("#00007F", "blue", "#007FFF", "cyan", 
      "white", "yellow", "#FF7F00", "red", "#7F0000"))(255)),
    na.value="transparent", values=col_val, 
    limits=c(min(bioclim_ewembi_1995_landonly$bio1)-2, 
             max(bioclim_ewembi_1995_landonly$bio1)+2)) + 
  coord_sf(expand=F, 
           xlim=c(min(bioclim_ewembi_1995_landonly$x), 
                  max(bioclim_ewembi_1995_landonly$x)), 
           ylim=c(min(bioclim_ewembi_1995_landonly$y),
                  max(bioclim_ewembi_1995_landonly$y)), 
           ndiscr=0) + theme_classic() + 
  theme(axis.title = element_blank(), axis.line = element_blank(),
        axis.ticks = element_blank(), axis.text = element_blank(),
        plot.background = element_rect(fill = "transparent"), 
        legend.background = element_rect(fill = "transparent"), 
        legend.box.background = element_rect(fill = "transparent", colour=NA))
```

**RCP2.6 - 2080**

```{r bio1_2080_rcp26, fig.width=9, fig.height=4}
data("bioclim_gfdl-esm2m_rcp26_2080_landonly")
data("bioclim_hadgem2-es_rcp26_2080_landonly")
data("bioclim_ipsl-cm5a-lr_rcp26_2080_landonly")
data("bioclim_miroc5_rcp26_2080_landonly")

bioclim_rcp26_2080_landonly <- bind_rows(`bioclim_gfdl-esm2m_rcp26_2080_landonly`, 
                                         `bioclim_hadgem2-es_rcp26_2080_landonly`, 
                                         `bioclim_ipsl-cm5a-lr_rcp26_2080_landonly`, 
                                         `bioclim_miroc5_rcp26_2080_landonly`) %>% 
  select(x,y,bio1) %>% group_by(x,y) %>% summarise(bio1=mean(bio1, na.rm=T))
col_val <- scales::rescale(unique(c(seq(min(bioclim_rcp26_2080_landonly$bio1), 0, length=5),
                                    seq(0, max(bioclim_rcp26_2080_landonly$bio1), length=5))))

ggplot() + geom_tile(data=bioclim_rcp26_2080_landonly, aes(x=x, y=y, fill=bio1)) + 
  geom_sf(data=outline, fill="transparent", colour="black") + 
  scale_fill_gradientn(name="tmean (°C)", colours=rev(colorRampPalette(
    c("#00007F", "blue", "#007FFF", "cyan", 
      "white", "yellow", "#FF7F00", "red", "#7F0000"))(255)),
    na.value="transparent", values=col_val, 
    limits=c(min(bioclim_rcp26_2080_landonly$bio1)-2, 
             max(bioclim_rcp26_2080_landonly$bio1)+2)) + 
  coord_sf(expand=F, 
           xlim=c(min(bioclim_rcp26_2080_landonly$x), 
                  max(bioclim_rcp26_2080_landonly$x)), 
           ylim=c(min(bioclim_rcp26_2080_landonly$y),
                  max(bioclim_rcp26_2080_landonly$y)), 
           ndiscr=0) + theme_classic() + 
  theme(axis.title = element_blank(), axis.line = element_blank(),
        axis.ticks = element_blank(), axis.text = element_blank(),
        plot.background = element_rect(fill = "transparent"), 
        legend.background = element_rect(fill = "transparent"), 
        legend.box.background = element_rect(fill = "transparent", colour=NA))
```

**RCP6.0 - 2080**

```{r bio1_2080_rcp60, fig.width=9, fig.height=4}
data("bioclim_gfdl-esm2m_rcp60_2080_landonly")
data("bioclim_hadgem2-es_rcp60_2080_landonly")
data("bioclim_ipsl-cm5a-lr_rcp60_2080_landonly")
data("bioclim_miroc5_rcp60_2080_landonly")

bioclim_rcp60_2080_landonly <- bind_rows(`bioclim_gfdl-esm2m_rcp60_2080_landonly`, 
                                         `bioclim_hadgem2-es_rcp60_2080_landonly`, 
                                         `bioclim_ipsl-cm5a-lr_rcp60_2080_landonly`,
                                         `bioclim_miroc5_rcp60_2080_landonly`) %>% 
  select(x,y,bio1) %>% group_by(x,y) %>% summarise(bio1=mean(bio1, na.rm=T))
col_val <- scales::rescale(unique(c(seq(min(bioclim_rcp60_2080_landonly$bio1), 0, length=5),
                                    seq(0, max(bioclim_rcp60_2080_landonly$bio1), length=5))))

ggplot() + geom_tile(data=bioclim_rcp60_2080_landonly, aes(x=x, y=y, fill=bio1)) + 
  geom_sf(data=outline, fill="transparent", colour="black") + 
  scale_fill_gradientn(name="tmean (°C)", colours=rev(colorRampPalette(
    c("#00007F", "blue", "#007FFF", "cyan", 
      "white", "yellow", "#FF7F00", "red", "#7F0000"))(255)),
    na.value="transparent", values=col_val, 
    limits=c(min(bioclim_rcp60_2080_landonly$bio1)-2, 
             max(bioclim_rcp60_2080_landonly$bio1)+2)) + 
  coord_sf(expand=F, 
           xlim=c(min(bioclim_rcp60_2080_landonly$x), 
                  max(bioclim_rcp60_2080_landonly$x)), 
           ylim=c(min(bioclim_rcp60_2080_landonly$y),
                  max(bioclim_rcp60_2080_landonly$y)), 
           ndiscr=0) + theme_classic() + 
  theme(axis.title = element_blank(), axis.line = element_blank(),
        axis.ticks = element_blank(), axis.text = element_blank(),
        plot.background = element_rect(fill = "transparent"), 
        legend.background = element_rect(fill = "transparent"), 
        legend.box.background = element_rect(fill = "transparent", colour=NA))
```

**RCP8.5 - 2080**

```{r bio1_2080_rcp85, fig.width=9, fig.height=4}
data("bioclim_gfdl-esm2m_rcp85_2080_landonly")
data("bioclim_hadgem2-es_rcp85_2080_landonly")
data("bioclim_ipsl-cm5a-lr_rcp85_2080_landonly")
data("bioclim_miroc5_rcp85_2080_landonly")

bioclim_rcp85_2080_landonly <- bind_rows(`bioclim_gfdl-esm2m_rcp85_2080_landonly`, 
                                         `bioclim_hadgem2-es_rcp85_2080_landonly`, 
                                         `bioclim_ipsl-cm5a-lr_rcp85_2080_landonly`,
                                         `bioclim_miroc5_rcp85_2080_landonly`) %>% 
  select(x,y,bio1) %>% group_by(x,y) %>% summarise(bio1=mean(bio1, na.rm=T))
col_val <- scales::rescale(unique(c(seq(min(bioclim_rcp85_2080_landonly$bio1), 0, length=5),
                                    seq(0, max(bioclim_rcp85_2080_landonly$bio1), length=5))))

ggplot() + geom_tile(data=bioclim_rcp85_2080_landonly, aes(x=x, y=y, fill=bio1)) + 
  geom_sf(data=outline, fill="transparent", colour="black") + 
  scale_fill_gradientn(name="tmean (°C)", colours=rev(colorRampPalette(
    c("#00007F", "blue", "#007FFF", "cyan", 
      "white", "yellow", "#FF7F00", "red", "#7F0000"))(255)),
    na.value="transparent", values=col_val, 
    limits=c(min(bioclim_rcp85_2080_landonly$bio1)-2, 
             max(bioclim_rcp85_2080_landonly$bio1)+2)) + 
  coord_sf(expand=F, 
           xlim=c(min(bioclim_rcp85_2080_landonly$x), 
                  max(bioclim_rcp85_2080_landonly$x)), 
           ylim=c(min(bioclim_rcp85_2080_landonly$y),
                  max(bioclim_rcp85_2080_landonly$y)), 
           ndiscr=0) + theme_classic() + 
  theme(axis.title = element_blank(), axis.line = element_blank(),
        axis.ticks = element_blank(), axis.text = element_blank(),
        plot.background = element_rect(fill = "transparent"), 
        legend.background = element_rect(fill = "transparent"), 
        legend.box.background = element_rect(fill = "transparent", colour=NA))
```

#### ISIMIP3b & GSWP3-W5E5

Current and future bioclimatic data for five 30-yr periods (1995, 2000, 2005, 2050, 2080) was derived from the GSWP3-W5E5 and ISIMIP3b data (https://esg.pik-potsdam.de/search/isimip/?project=ISIMIP3b&amp;product=input&amp;dataset_type=Climate+atmosphere+simulated) and is included in this package. 

**GSWP3_W5E5 - 2005**

```{r bio1_2005, fig.width=9, fig.height=4}
data("bioclim_gswp3-w5e5_obsclim_2005_landonly")

library(dplyr); library(sf); library(ggplot2)
data(outline, package="ggmap2")
outline <- sf::st_as_sf(outline)
col_val <- scales::rescale(unique(c(seq(min(`bioclim_gswp3-w5e5_obsclim_2005_landonly`$bio1), 0, length=5),
                                    seq(0, max(`bioclim_gswp3-w5e5_obsclim_2005_landonly`$bio1), length=5))))

`bioclim_gswp3-w5e5_obsclim_2005_landonly` %>% select(x,y,bio1) %>% 
  ggplot() + geom_tile(aes(x=x, y=y, fill=bio1)) + 
  geom_sf(data=outline, fill="transparent", colour="black") + 
  scale_fill_gradientn(name="tmean (°C)", colours=rev(colorRampPalette(
    c("#00007F", "blue", "#007FFF", "cyan", 
      "white", "yellow", "#FF7F00", "red", "#7F0000"))(255)),
    na.value="transparent", values=col_val, 
    limits=c(min(`bioclim_gswp3-w5e5_obsclim_2005_landonly`$bio1)-2, 
             max(`bioclim_gswp3-w5e5_obsclim_2005_landonly`$bio1)+2)) + 
  coord_sf(expand=F, 
           xlim=c(min(`bioclim_gswp3-w5e5_obsclim_2005_landonly`$x), 
                  max(`bioclim_gswp3-w5e5_obsclim_2005_landonly`$x)), 
           ylim=c(min(`bioclim_gswp3-w5e5_obsclim_2005_landonly`$y),
                  max(`bioclim_gswp3-w5e5_obsclim_2005_landonly`$y)), 
           ndiscr=0) + theme_classic() + 
  theme(axis.title = element_blank(), axis.line = element_blank(),
        axis.ticks = element_blank(), axis.text = element_blank(),
        plot.background = element_rect(fill = "transparent"), 
        legend.background = element_rect(fill = "transparent"), 
        legend.box.background = element_rect(fill = "transparent", colour=NA))
```

**SSP126 - 2080**

```{r bio1_2080_ssp126, fig.width=9, fig.height=4}
data("bioclim_gfdl-esm4_ssp126_2080_landonly")
data("bioclim_ipsl-cm6a-lr_ssp126_2080_landonly")
data("bioclim_mpi-esm1-2-hr_ssp126_2080_landonly")
data("bioclim_mri-esm2-0_ssp126_2080_landonly")
data("bioclim_ukesm1-0-ll_ssp126_2080_landonly")

bioclim_ssp126_2080_landonly <- bind_rows(`bioclim_gfdl-esm4_ssp126_2080_landonly`, 
                                         `bioclim_ipsl-cm6a-lr_ssp126_2080_landonly`, 
                                         `bioclim_mpi-esm1-2-hr_ssp126_2080_landonly`, 
                                         `bioclim_mri-esm2-0_ssp126_2080_landonly`,
                                         `bioclim_ukesm1-0-ll_ssp126_2080_landonly`) %>% 
  select(x,y,bio1) %>% group_by(x,y) %>% summarise(bio1=mean(bio1, na.rm=T))
col_val <- scales::rescale(unique(c(seq(min(bioclim_ssp126_2080_landonly$bio1), 0, length=5),
                                    seq(0, max(bioclim_ssp126_2080_landonly$bio1), length=5))))

ggplot() + geom_tile(data=bioclim_ssp126_2080_landonly, aes(x=x, y=y, fill=bio1)) + 
  geom_sf(data=outline, fill="transparent", colour="black") + 
  scale_fill_gradientn(name="tmean (°C)", colours=rev(colorRampPalette(
    c("#00007F", "blue", "#007FFF", "cyan", 
      "white", "yellow", "#FF7F00", "red", "#7F0000"))(255)),
    na.value="transparent", values=col_val, 
    limits=c(min(bioclim_ssp126_2080_landonly$bio1)-2, 
             max(bioclim_ssp126_2080_landonly$bio1)+2)) + 
  coord_sf(expand=F, 
           xlim=c(min(bioclim_ssp126_2080_landonly$x), 
                  max(bioclim_ssp126_2080_landonly$x)), 
           ylim=c(min(bioclim_ssp126_2080_landonly$y),
                  max(bioclim_ssp126_2080_landonly$y)), 
           ndiscr=0) + theme_classic() + 
  theme(axis.title = element_blank(), axis.line = element_blank(),
        axis.ticks = element_blank(), axis.text = element_blank(),
        plot.background = element_rect(fill = "transparent"), 
        legend.background = element_rect(fill = "transparent"), 
        legend.box.background = element_rect(fill = "transparent", colour=NA))
```

**SSP370 - 2080**

```{r bio1_2080_ssp370, fig.width=9, fig.height=4}
data("bioclim_gfdl-esm4_ssp370_2080_landonly")
data("bioclim_ipsl-cm6a-lr_ssp370_2080_landonly")
data("bioclim_mpi-esm1-2-hr_ssp370_2080_landonly")
data("bioclim_mri-esm2-0_ssp370_2080_landonly")
data("bioclim_ukesm1-0-ll_ssp370_2080_landonly")

bioclim_ssp370_2080_landonly <- bind_rows(`bioclim_gfdl-esm4_ssp370_2080_landonly`, 
                                         `bioclim_ipsl-cm6a-lr_ssp370_2080_landonly`, 
                                         `bioclim_mpi-esm1-2-hr_ssp370_2080_landonly`, 
                                         `bioclim_mri-esm2-0_ssp370_2080_landonly`,
                                         `bioclim_ukesm1-0-ll_ssp370_2080_landonly`) %>% 
  select(x,y,bio1) %>% group_by(x,y) %>% summarise(bio1=mean(bio1, na.rm=T))
col_val <- scales::rescale(unique(c(seq(min(bioclim_ssp370_2080_landonly$bio1), 0, length=5),
                                    seq(0, max(bioclim_ssp370_2080_landonly$bio1), length=5))))

ggplot() + geom_tile(data=bioclim_ssp370_2080_landonly, aes(x=x, y=y, fill=bio1)) + 
  geom_sf(data=outline, fill="transparent", colour="black") + 
  scale_fill_gradientn(name="tmean (°C)", colours=rev(colorRampPalette(
    c("#00007F", "blue", "#007FFF", "cyan", 
      "white", "yellow", "#FF7F00", "red", "#7F0000"))(255)),
    na.value="transparent", values=col_val, 
    limits=c(min(bioclim_ssp370_2080_landonly$bio1)-2, 
             max(bioclim_ssp370_2080_landonly$bio1)+2)) + 
  coord_sf(expand=F, 
           xlim=c(min(bioclim_ssp370_2080_landonly$x), 
                  max(bioclim_ssp370_2080_landonly$x)), 
           ylim=c(min(bioclim_ssp370_2080_landonly$y),
                  max(bioclim_ssp370_2080_landonly$y)), 
           ndiscr=0) + theme_classic() + 
  theme(axis.title = element_blank(), axis.line = element_blank(),
        axis.ticks = element_blank(), axis.text = element_blank(),
        plot.background = element_rect(fill = "transparent"), 
        legend.background = element_rect(fill = "transparent"), 
        legend.box.background = element_rect(fill = "transparent", colour=NA))
```

**SSP585 - 2080**

```{r bio1_2080_ssp585, fig.width=9, fig.height=4}
data("bioclim_gfdl-esm4_ssp585_2080_landonly")
data("bioclim_ipsl-cm6a-lr_ssp585_2080_landonly")
data("bioclim_mpi-esm1-2-hr_ssp585_2080_landonly")
data("bioclim_mri-esm2-0_ssp585_2080_landonly")
data("bioclim_ukesm1-0-ll_ssp585_2080_landonly")

bioclim_ssp585_2080_landonly <- bind_rows(`bioclim_gfdl-esm4_ssp585_2080_landonly`, 
                                         `bioclim_ipsl-cm6a-lr_ssp585_2080_landonly`, 
                                         `bioclim_mpi-esm1-2-hr_ssp585_2080_landonly`, 
                                         `bioclim_mri-esm2-0_ssp585_2080_landonly`,
                                         `bioclim_ukesm1-0-ll_ssp585_2080_landonly`) %>% 
  select(x,y,bio1) %>% group_by(x,y) %>% summarise(bio1=mean(bio1, na.rm=T))
col_val <- scales::rescale(unique(c(seq(min(bioclim_ssp585_2080_landonly$bio1), 0, length=5),
                                    seq(0, max(bioclim_ssp585_2080_landonly$bio1), length=5))))

ggplot() + geom_tile(data=bioclim_ssp585_2080_landonly, aes(x=x, y=y, fill=bio1)) + 
  geom_sf(data=outline, fill="transparent", colour="black") + 
  scale_fill_gradientn(name="tmean (°C)", colours=rev(colorRampPalette(
    c("#00007F", "blue", "#007FFF", "cyan", 
      "white", "yellow", "#FF7F00", "red", "#7F0000"))(255)),
    na.value="transparent", values=col_val, 
    limits=c(min(bioclim_ssp585_2080_landonly$bio1)-2, 
             max(bioclim_ssp585_2080_landonly$bio1)+2)) + 
  coord_sf(expand=F, 
           xlim=c(min(bioclim_ssp585_2080_landonly$x), 
                  max(bioclim_ssp585_2080_landonly$x)), 
           ylim=c(min(bioclim_ssp585_2080_landonly$y),
                  max(bioclim_ssp585_2080_landonly$y)), 
           ndiscr=0) + theme_classic() + 
  theme(axis.title = element_blank(), axis.line = element_blank(),
        axis.ticks = element_blank(), axis.text = element_blank(),
        plot.background = element_rect(fill = "transparent"), 
        legend.background = element_rect(fill = "transparent"), 
        legend.box.background = element_rect(fill = "transparent", colour=NA))
```
