---
title: "Variable correlation & climate scenario comparison"
author: "RS-eco"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Variable correlation & climate scenario comparison}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width=14, fig.height=8, warning=FALSE, comment=NA, message=FALSE, eval=F)
```

## Bioclim correlation

Read in the bioclimatic variables, which are included in the package. **Note:** The code for creating the bioclimg files (bioclim.R) is included in the data-raw folder on Github. <!-- Add direct link -->

```{r}
#Load rISIMIP package
library(rISIMIP)

#List data files
d <- data(package="rISIMIP")
bioclim_files <- d$results[,"Item"][grep(x=d$results[,"Item"], pattern="bioclim")]

#Load data as raster
bioclim_ref <- lapply(bioclim_files, function(x) raster::rasterFromXYZ(get(data(list=x))))
bioclim_ref <- lapply(bioclim_ref, function(x){
  names(x) <- paste0("bio", 1:19)
  return(x)
})
```

## Check for correlation using variance inflation factors

VIF can be used to detect collinearity (Strong correlation between two or more predictor variables). Collinearity causes instability in parameter estimation in regression-type models. The VIF is based on the square of the multiple correlation coefficient resulting from regressing a predictor variable against all other predictor variables. If a variable has a strong linear relationship with at least one other variables, the correlation coefficient would be close to 1, and VIF for that variable would be large. A VIF greater than 10 is a signal that the model has a collinearity problem.

Load usdm R-package from Babak Naimi

```{r}
#install.packages("usdm")
library(usdm)
```

Vif function calculates this statistic for all variables in x.

```{r}
vif_bio <- lapply(bioclim_ref, FUN=function(x) vif(x, maxobservations=201600))
vif_bio <- lapply(vif_bio, function(x){
  x$Variables <- paste0("bio", 1:19)
  return(x)
  })
vif_bio <- Reduce(function(dtf1,dtf2) dplyr::full_join(dtf1, dtf2, by="Variables"), vif_bio)
colnames(vif_bio) <- c("Variables", sub("bioclim_", "", sub("_landonly.csv.xz", "", basename(bioclim_files))))
knitr::kable(vif_bio)
```

vifcor and vifstep uses two different strategy to exclude highly collinear variable through a stepwise procedure. 
   
vifcor, first find a pair of variables which has the maximum linear correlation (greater than th), and exclude one of them which has greater VIF. The procedure is repeated untill no variable with a high corrrelation coefficient (greater than threshold) with other variables remains. 

```{r}
vifcor_bio <- lapply(1:9, FUN=function(x) vifcor(bioclim_ref[[x]], 
                                                 th=0.7, maxobservations=300000))
vifcor_bio <- lapply(vifcor_bio, FUN=function(x) x@results)
vifcor_bio <- Reduce(function(dtf1,dtf2) dplyr::full_join(dtf1, dtf2, by="Variables"), vifcor_bio)
colnames(vifcor_bio) <- c("Variables", sub("bioclim_", "", sub("_landonly.csv.xz", "", basename(bioclim_files))))
knitr::kable(vifcor_bio)
```

vifstep calculate VIF for all variables, exclude one with highest VIF (greater than threshold), 
repeat the procedure untill no variables with VIF greater than the threshold remains.

```{r}
vifstep_bio <- lapply(1:9, FUN=function(x) vifstep(bioclim_ref[[x]], 
                                                   th=6, maxobservations=70000)) 
vifstep_bio <- lapply(vifstep_bio, FUN=function(x) x@results)
vifstep_bio <- Reduce(function(dtf1,dtf2) dplyr::full_join(dtf1, dtf2, by="Variables"), vifstep_bio)
colnames(vifstep_bio) <- c("Variables", sub("bioclim_", "", sub("_landonly.csv.xz", "", basename(bioclim_files))))
knitr::kable(vifstep_bio)
```

## Climate scenario comparison

```{r}
# Read Bioclim data files as one data.frame
library(dplyr)
bio_df <- lapply(bioclim_files, function(x){
  data <- get(data(list=x))
  data$Year <-  strsplit(basename(x), split="_")[[1]][4]
  data$Year[data$Year == "landonly"] <- 1995
  data$Model <- strsplit(basename(x), split="_")[[1]][2]
  data$RCP <- strsplit(basename(x), split="_")[[1]][3]
  data$RCP[data$RCP == 1995] <- NA
  return(data)
})
bio_df <- dplyr::bind_rows(bio_df)

# Create summary table
bio_mean <- bio_df %>% dplyr::select(-c(x,y, bio1, bio2, bio3, bio6, bio7, bio8, bio9, bio10, bio11, bio13, bio14, bio16, bio17)) %>% 
  group_by(Year, Model, RCP) %>% summarise_all(funs(mean, sd))
bio_mean[,-c(1:3)] <- round(bio_mean[,-c(1:3)], digits=2)
bio_mean <- bio_mean %>% tidyr::unite("bio4", "bio4_mean", "bio4_sd",sep=" ± ") %>% tidyr::unite("bio5", "bio5_mean", "bio5_sd",sep=" ± ") %>% tidyr::unite("bio12", "bio12_mean", "bio12_sd",sep=" ± ") %>% tidyr::unite("bio15", "bio15_mean", "bio15_sd",sep=" ± ") %>% tidyr::unite("bio18", "bio18_mean", "bio18_sd",sep=" ± ") %>% tidyr::unite("bio19", "bio19_mean", "bio19_sd",sep=" ± ")
```

**Table 1.** Mean ± SD of global bioclim values per model.

```{r, results="asis"}
knitr::kable(bio_mean)
```

## Bioclim & land-use correlation

**Note:** The code for creating the landuse files (landuse.R) is included in the data-raw folder on Github. <!-- Add direct link -->

```{r}
# Read Landuse data
#List data files
d <- data(package="rISIMIP")
landuse <- d$results[,"Item"][grep(x=d$results[,"Item"], pattern="landuse-totals_")][2:17]
landuse_df <- lapply(landuse, function(x){
  data <- get(data(list=x))
  data$Year <-  strsplit(strsplit(basename(x), split="_")[[1]][4], split="[.]")[[1]][1]
  data$Model <- strsplit(basename(x), split="_")[[1]][2]
  data$RCP <- strsplit(basename(x), split="_")[[1]][1]
  return(data)
})
landuse_df <- do.call(plyr::rbind.fill, landuse_df)
#head(landuse_df)

# Create summary table
landuse_df$biofuel_cropland_irrigated[is.na(landuse_df$biofuel_cropland_irrigated)] <- 0
landuse_df$biofuel_cropland_rainfed[is.na(landuse_df$biofuel_cropland_rainfed)] <- 0
landuse_mean <- landuse_df %>% dplyr::select(-c(x,y)) %>% group_by(Year, Model, RCP) %>% summarise_all(funs(mean, sd), na.rm=TRUE)
landuse_mean[,-c(1:3)] <- round(landuse_mean[,-c(1:3)], digits=2)
landuse_mean <- landuse_mean %>% tidyr::unite("cropland_irrigated", "cropland_irrigated_mean", "cropland_irrigated_sd",sep=" ± ") %>% tidyr::unite("cropland_rainfed", "cropland_rainfed_mean", "cropland_rainfed_sd", sep=" ± ") %>% tidyr::unite("cropland_total", "cropland_total_mean", "cropland_total_sd",sep=" ± ") %>% tidyr::unite("pastures", "pastures_mean", "pastures_sd",sep=" ± ") %>% tidyr::unite("biofuel_cropland_irrigated", "biofuel_cropland_irrigated_mean", "biofuel_cropland_irrigated_sd",sep=" ± ") %>% tidyr::unite("biofuel_cropland_rainfed", "biofuel_cropland_rainfed_mean", "biofuel_cropland_rainfed_sd",sep=" ± ")
```

**Table 2.** Mean ± SD of global landuse values per model.

```{r, results="asis"}
knitr::kable(landuse_mean)
```

## Correlation between bioclim and land-use data

```{r}
library(corrplot)

# Climate
clim1995 <- get(data("bioclim_ewembi_1995_landonly"))
CorD <- cor(clim1995[,c(3,6,7,8,12,13,14,17,20,21)])

clim2080 <- get(data("bioclim_gfdl-esm2m_rcp26_2080_landonly"))
CorD <- cor(clim2080[,c(3,6,7,8,12,13,14,17,20,21)])

CorForP <- abs(CorD)
corrplot(CorD, method="circle", bg = "white",addgrid.col = "gray10", 
         tl.col = "black",tl.cex = 0.8, p.mat = CorForP, sig.level = 0.7)

# Landuse
landuse1995 <- get(data("rcp26_gfdl-esm2m_landuse-totals_1995"))
CorD <- cor(landuse1995[,c(3:ncol(landuse1995))])

landuse2080 <- get(data("rcp26_gfdl-esm2m_landuse-totals_2080"))
CorD <- cor(landuse2080[,c(3:ncol(landuse2080))])

CorForP <- abs(CorD)
corrplot(CorD, method="circle", bg = "white",addgrid.col = "gray10", tl.col = "black",
         tl.cex = 0.8, p.mat = CorForP, sig.level = 0.7)

# Climate and land-use
CorD <- left_join(clim2080, landuse2080, by=c("x", "y")) %>% 
  dplyr::select(-c(x, y, bio1, bio2, bio3, bio6, bio7, bio8, bio9, bio10, 
            bio11, bio13, bio14, bio16, bio17)) %>% cor(use="complete.obs")
CorForP <- abs(CorD)
corrplot(CorD, method="circle", bg = "white",addgrid.col = "gray10", tl.col = "black",
         tl.cex = 0.8, p.mat = CorForP, sig.level = 0.7)
```
