---
title: "Derive global landonly ISIMIP data"
author: "RS-eco"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Derive global landonly ISIMIP data}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, warning=F, comment=NA, message=F, eval=F)
```
  
## Data Setup
  
Install rISIMIP package
  
```{r, eval=FALSE}
# Install remotes if not previously installed
if(!"remotes" %in% installed.packages()[,"Package"]) install.packages("remotes")

# Install rISIMIP from Github if not previously installed
if(!"rISIMIP" %in% installed.packages()[,"Package"]) remotes::install_github("RS-eco/rISIMIP")
```
  
Load rISIMIP package
  
```{r}
library(rISIMIP)
```
  
First, we specify the file path, where the ISIMIP2b data is located.
  
**Note:** The following script requires that you already have the ISIMIP2b bias-corrected GCM_atmosphere data on your local harddrive.
  
```{r global_options}
# Specify path of file directory
#filedir <- "/media/matt/Data/Documents/Wissenschaft/Data/"
filedir <- "/work/bb0820/ISIMIP/"
```

## Summarise climate data

First, we need the `summariseNC` function from the `processNC` package.
Thus, we have to install the processNC package from Github.

```{r install_processNC}
# Install processNC from Github if not previously installed
if(!"processNC" %in% installed.packages()[,"Package"]) remotes::install_github("RS-eco/processNC")
```

Load processNC package

```{r load_processNC}
library(processNC)
```

We now list and summarise the climate data for the required time steps (2006-2035, 2036-2065, 2066-2095, 2086-2115, 2136-2165) for each variable and each model using global ISIMIP2b data.

## Certain years + 1.5 degree time slices

```{r}
#Timeframes
timeframe <- c("1845","1990","1995","2009","2010","2020","2026","2032","2048",
               "2050","2052", "2056","2080","2100","2150","2200","2250")
startyear <- c(1830,1976,1980,1995,1996,2006,2012,2018,2034,
               2036,2038,2042,2066,2086,2136,2186,2236)
endyear <- c(1859,2005,2009,2024,2025,2035,2041,2047,2063,
             2065,2067,2071,2095,2115,2165,2215,2265)
timeperiods <- data.frame(timeframe=timeframe, startyear=startyear,endyear=endyear)

#Climate variables
vars <- c("tasmin", "tasmax", "pr")

#Climate models
models <- c("GFDL-ESM2M", "HadGEM2-ES", "IPSL-CM5A-LR", "MIROC5")

#RCP scenarios
rcps <- c("rcp26", "rcp60")

#Create list of variable, climate model and time frame combination
var_mod_time <- expand.grid(var = vars, model = models, 
                            timeframe = timeframe, rcp = rcps)

# Add historical scenario
df <- expand.grid(rcp="historical", var=vars, model=models, timeframe = "1990")
var_mod_time <- rbind(df, var_mod_time)

# Add EWEMBI scenario
df <- expand.grid(rcp="1995", var=vars, model="EWEMBI", timeframe = "1995")
var_mod_time <- rbind(df, var_mod_time)

# Add piControl scenario
df <- expand.grid(rcp="piControl", var=vars, model=models, timeframe = "1845")
var_mod_time <- rbind(df, var_mod_time)
rm(vars, models, timeframe, rcps)

var_mod_time <- dplyr::left_join(var_mod_time, timeperiods, by="timeframe"); rm(timeperiods)
```

## Slow method

**Note:** Be aware that the following code chunk will take a very long time to run, even on a high-performance computer. A faster method is shown in the 2nd R code chunk below, however this approach requires the Climate Data Operators programme installed on your computer.

```{r, eval=F}
# Run summariseNC for all combinations
lapply(1:nrow(var_mod_time), FUN=function(x){
  files <- listISIMIP(path=filedir, var=var_mod_time$var[x], extent="global", 
                      model=var_mod_time$model[x], 
                      scenario=var_mod_time$rcp[x], 
                      startyear=var_mod_time$startyear[x], 
                      endyear=var_mod_time$endyear[x])
  filename1 <- paste0(filedir, "/ISIMIP2b/DerivedInputData/GCM/global/", 
                      var_mod_time$rcp[x], "/monthly_", 
                      var_mod_time$var[x], "_", var_mod_time$model[x], "_",
                      var_mod_time$rcp[x], "_", var_mod_time$timeframe[x], 
                      ".nc4")
  if(length(files)==4){
    if(!file.exists(filename1)){
      data_sub <- summariseNC(files=files, 
                              startdate=var_mod_time$startyear[x], 
                              enddate=var_mod_time$endyear[x], 
                              filename1=filename1, format="CDF", 
                              overwrite=FALSE)
    }
  }
  print(x/nrow(var_mod_time)*100)
})
#system("shutdown -s -f")
```

## Fast method using CDO

**Advice:** Run this directly on ISIMIP Server, saves you the download of the data.

```{r, eval=F}
#Turn var_mod_time into correct order
var_mod_time <- dplyr::arrange(var_mod_time, rcp, model, var)

#Create outfile list
var_mod_time$outfile <- sapply(1:nrow(var_mod_time), FUN=function(x){
  paste0(filedir, "/ISIMIP2b/", var_mod_time$var[x], "_day_", var_mod_time$model[x], "_", 
         var_mod_time$rcp[x], "_", floor(var_mod_time$startyear[x]/10)*10+1, "_", 
         ceiling(var_mod_time$endyear[x]/10)*10, ".nc4")
})
outfiles <- unique(var_mod_time$outfile)

#Create oufile2 list
var_mod_time$outfile2 <- sapply(1:nrow(var_mod_time), FUN=function(x){
  paste0(filedir, "/ISIMIP2b/DerivedInputData/GCM/global/", var_mod_time$rcp[x], "/monthly_", var_mod_time$var[x], "_", 
         var_mod_time$model[x], "_", var_mod_time$rcp[x], "_", var_mod_time$timeframe[x], ".nc4")
})

#Run mergeNC for all unique outfiles
lapply(outfiles, FUN=function(x){
  # Subset var_mod_time according to outfiles
  var_mod_sub <- var_mod_time[var_mod_time$outfile == x,]

  if(any(!file.exists(var_mod_sub$outfile2))){
    # List ISIMIP2b files
    files <- rISIMIP::listISIMIP(path=paste0(filedir, 
                                             "/ISIMIP2b/InputData/GCM/"), 
                                 var=var_mod_sub$var[1], 
                                 extent="global", model=var_mod_sub$model[1], 
                                 scenario=var_mod_sub$rcp[1], 
                                 startyear=var_mod_sub$startyear[1], endyear=var_mod_sub$endyear[1])
    files <- Filter(Negate(is.na), files)
    if(all(var_mod_sub$startyear >= 2006 & length(files)==4) | 
       all(var_mod_sub$startyear < 2006 & length(files) == 5) | 
       all(var_mod_sub$rcp == "piControl" & length(files) == 4) |
       all(var_mod_sub$rcp == "historical" & length(files) == 4)){
      if(!file.exists(x)){
        processNC::mergeNC(files, x)
      }
      # Run aggregateNC for all combinations
      lapply(1:nrow(var_mod_sub), FUN=function(y){
        if(!file.exists(var_mod_sub$outfile2[y])){
          processNC::aggregateNC(infile=x, outfile=var_mod_sub$outfile2[y], 
                                 var=var_mod_sub$var[y], 
                      startdate=var_mod_sub$startyear[y], 
                      enddate=var_mod_sub$endyear[y])
        }
      })
      file.remove(x)
    } else{
      print(files)
    }
  }
})
#system("shutdown -s -f")
```

## Transform climate data

First, we make a list of the newly created files:

```{r filelist}
#List all tasmin and tasmax files
tmin_files <- list.files(
  paste0(getwd(), "/ISIMIP2b/DerivedInputData/GCM/global/"),
  pattern="monthly_tasmin_.*\\.nc4", full.names=T, recursive=T)
tmax_files <- list.files(
  paste0(getwd(), "/ISIMIP2b/DerivedInputData/GCM/global/"), 
  pattern="monthly_tasmax_.*\\.nc4", full.names=T, recursive=T)

# List precipitation files
prec_files <-  list.files(
  paste0(getwd(), "/ISIMIP2b/DerivedInputData/GCM/global/"),
  pattern="monthly_pr_.*\\.nc4", full.names=T, recursive=T)

# Select certain years
years <- unique(var_mod_time$timeframe)[c(2,4:13)]

tmin_files <- unlist(lapply(years, function(x) tmin_files[grep(tmin_files, pattern=x)]))
tmax_files <- unlist(lapply(years, function(x) tmax_files[grep(tmax_files, pattern=x)]))
prec_files <- unlist(lapply(years, function(x) prec_files[grep(prec_files, pattern=x)]))
  
# Check tmin, tmax and prec files are identical, 
#not just same number of files
```

### Subset data

We want to get global data, but for land only:

```{r landonly, eval=F}
# Read landonly mask
data("landseamask_generic", package="rISIMIP")

# Mask data by landonly mask
library(raster)
tmin_lo <- lapply(tmin_files, FUN=function(x) mask(stack(x), landseamask_generic))
tmax_lo <- lapply(tmax_files, FUN=function(x) mask(stack(x), landseamask_generic))
prec_lo <- lapply(prec_files, FUN=function(x) mask(stack(x), landseamask_generic))

plot(tmin_lo[[5]])
plot(tmax_lo[[5]])
plot(prec_lo[[5]])

# Merge lists into one list
data_lo <- c(tmin_lo, tmax_lo, prec_lo)
data_files <- c(tmin_files, tmax_files, prec_files)

# Save to file in landonly subfolder
mapply(FUN=function(x,y){
  filename <- sub(".nc4", "_landonly.nc", gsub("global", "landonly", y))
  if(!file.exists(sub(".nc4", ".nc", filename))){
    x <- stack(x)
    #x <- as.data.frame(rasterToPoints(x))
    #colnames(x) <- c("x", "y", month.abb)
    raster::writeRaster(x, filename=filename, format="CDF",
                        xname="lon", yname="lat", zname="time", 
                        zunit="years since 1661-1-1 00:00:00",
                        force_v4=TRUE, compression=9)
  }
}, x=data_lo, y=data_files)

# Change file ending to .nc4
lapply(data_files, function(x){
  file.rename(sub(".nc4", "_landonly.nc", 
                    gsub("global", "landonly", x)), sub(".nc4", "_landonly.nc4", 
                    gsub("global", "landonly", x)))
})
```

#### Change units

The climate data comes in non-standard units. Temperature is in Kelvin and needs to be converted to degree Celsius, while precipitation was originally in kg m-2 s-1 and needs to be converted to kg m-2 day-1, which equals mm per day.

```{r units, eval=F}
#Turn climate data into right units (degree Celsius and mm)

#Convert temperature from Kelvin to degree Celsius
tmin_lo <- lapply(tmin_lo, FUN=function(x){
  raster::calc(x, fun=function(x){x-273.15})
})
tmax_lo <- lapply(tmax_lo, FUN=function(x){
  raster::calc(x, fun=function(x){x-273.15})
})

# Convert precipitation from kg m-2 s-1 to kg m-2 day-1
prec_lo <- lapply(prec_lo, FUN=function(x){
  raster::calc(x, fun=function(x){x*86400})
})

plot(tmin_lo[[1]][[1]])
plot(tmax_lo[[1]][[1]])
plot(prec_lo[[1]][[1]])
```

### Calculate Bioclim variables

From the resulting layers, we can now calculate the Bioclimatic variables, using the biovars function in the dismo package.

```{r bioclim, eval=F}
library(raster)

# Create list with bioclim names
bioclim_names <- gsub(x = prec_files, pattern = "\\monthly_pr", 
                      replacement = "bioclim")
bioclim_names <- sub(".nc4", "_landonly.nc", 
                     gsub("global", "landonly", bioclim_names))

# Calculate bioclim variables for all models and time frames and save to file
bioclim <- mapply(FUN=function(x,y,z,name){
  if(!file.exists(sub(".nc4", ".nc", name))){
    bio <- dismo::biovars(tmin=x, tmax=y, prec=z)
    raster::writeRaster(bio, filename=sub(".nc4", ".nc", name), format="CDF", 
                        xname="lon", yname="lat", zname="time", 
                        zunit="years since 1661-1-1 00:00:00",
                        force_v4=TRUE, compression=9)
  }
}, x=tmin_lo, y=tmax_lo, z=prec_lo, name=bioclim_names)

# Change file name from .nc to .nc4
bioclim_files <- list.files(
  paste0(getwd(), "/ISIMIP2b/DerivedInputData/GCM/landonly"),
  pattern="bioclim_.*\\.nc", full.names=T, recursive=T)
#lapply(bioclim_files, function(x) raster::extension(x, ".nc4"))
file.rename(bioclim_files, sub(".nc", ".nc4", bioclim_files))
```

Now, we can list the previously calculated bioclim files

```{r}
bioclim_files <- list.files(
  paste0(filedir, "/ISIMIP2b/DerivedInputData/GCM/landonly"),
  pattern="bioclim_.*\\.nc4", full.names=T, recursive=T)
bioclim <- lapply(bioclim_files, raster::stack)

# Need to implement bioclim_names here!
#bioclim_names 

# List internal bioclim files
#(bioclim_files <- list.files("data", pattern="bioclim_.*\\landonly.rda", 
#                            full.names=T, recursive=T))
```

## Data visualisation

### ggmap2 package

For plotting our data with the `createMap` function, we need to install the `ggmap2` package.

```{r, eval=FALSE}
# Install ggmap2 package from Github
devtools::install_github("RS-eco/ggmap2")
```

Load ggmap2 package

```{r}
library(ggmap2)
```

### Historic (1985) global landonly temperature

```{r, eval=FALSE}
# Read tas landonly data files
tas <- lapply(
  list.files(paste0(filedir, "/ISIMIP2b/DerivedInputData/GCM/landonly/historical"), 
             pattern="monthly_tas_.*\\.nc", full.names=TRUE), stack)

# Create Map
createMap(tas[[1]], name="tas", subnames=month.abb, 
          split=FALSE, ncol=3, width=12, height=8, units="in", dpi=100)
```

### Plot Bioclim data

```{r, bioclim_2050, eval=FALSE}
# Read Bioclim data files
bioclim_2050 <- lapply(list.files(paste0(filedir, "ISIMIP2b/ProcessedData/bioclim"),
                                  pattern="rcp_2050.nc", full.names=TRUE), stack)

# Create Map of bioclim data
bio04_2050 <- stack(lapply(bioclim_2050, function(x) x[[4]]))
createMap(bioclim[[20]][[4]], name="Bio04", split=FALSE, ncol=2, width=8, height=12, units="in", filename=NA, dpi=100)
```

### Plot all climate data and save to file

```{r eval=FALSE, echo=FALSE}
# Create maps for all variables, all models, all scenarios and all timeframes
vars <- c("hurs", "huss", "tas", "tasmin", "tasmax", "pr")
#scenarios <- c("historical", "rcp26", "rcp60")

files <- sapply(vars, function(x) list.files(paste0(filedir, "/ISIMIP2b/DerivedInputData/GCM/landonly/"), pattern=paste0("monthly_", x, "_.*\\.nc"), full.names=T, recursive=T))
files <- do.call("c", files)

mapply(FUN=function(data,name) createMap(data=data, name=x, subnames=month.abb, split=FALSE, ncol=3, width=12, height=8, units="in", filename=name, dpi=100), data=lapply(files, stack), name=sapply(files, FUN=function(x) paste0("figures/", sub(".grd", ".tiff", basename(x)))))
```

```{r, eval=FALSE}
library(readr)
tmax_files <- list.files("/media/matt/Data/Documents/Wissenschaft/Data/ISIMIP2b/DerivedInputData/30yr_climate/", pattern="monthly_tasmax_.*\\.nc", full.names=TRUE, recursive=T)
for(i in 1:length(tmax_files)){
  tasmax <- stack(tmax_files[i])
  tasmax_df <- as.data.frame(rasterToPoints(tasmax))
  colnames(tasmax_df) <- c("x", "y", month.abb)
  ggplot() + geom_raster(data=tasmax_df, aes(x=x, y=y, fill=Jul)) + 
    scale_fill_gradientn(colours=
                           colorRampPalette(c("#00007F", "blue", "#007FFF", "cyan",
                                              "#7FFF7F", "yellow", 
                                              "#FF7F00", "red", "#7F0000"))(255))
  ggsave(sub(pattern=".grd", ".png", tmax_files[i]))
  readr::write_csv(tasmax_df, sub(pattern=".grd", ".csv", tmax_files[i]))
}
file.remove(tmax_files)
file.remove(sub(pattern=".grd", ".gri", tmax_files))

bioclim_files <- list.files("/media/matt/Data/Documents/Wissenschaft/Data/ISIMIP2b/DerivedInputData/30yr_climate/rcp85",
                            pattern="bioclim_.*\\.nc4", full.names=TRUE, recursive=T)
for(i in 1:length(bioclim_files)){
  bioclim <- stack(bioclim_files[i])
  bioclim_df <- as.data.frame(rasterToPoints(bioclim))
    colnames(bioclim_df) <- c("x", "y", paste0("bio", 1:19))
    ggplot() + geom_tile(data=bioclim_df, aes(x=x, y=y, fill=bio5)) + 
    scale_fill_gradientn(colours=
                           colorRampPalette(c("#00007F", "blue", "#007FFF", "cyan",
                                              "#7FFF7F", "yellow", 
                                              "#FF7F00", "red", "#7F0000"))(255)) + 
      geom_sf()
  ggsave(sub(pattern=".nc4", ".png", bioclim_files[i]))
  readr::write_csv(bioclim_df, sub(pattern=".nc4", ".csv.xz", bioclim_files[i]))
}
file.remove(bioclim_files)
```

## Plot change in climate

```{r}
bioclim_1995 <- read.csv(list.files(
  paste0(filedir, "/ISIMIP2b/DerivedInputData/GCM/landonly"),
  pattern="bioclim_EWEMBI.*\\.nc", 
  full.names=T, recursive=T))[,c("x", "y", "bio4", "bio5", "bio12", "bio15", "bio18", "bio19")]
bioclim_1995$year <- 1995
bioclim_1995 <- tidyr::gather(bioclim_1995, var, value, -c(x,y,year))
bioclim_1995 <- tidyr::spread(bioclim_1995, year, value)

bioclim_fut <- c(list.files(
  paste0(filedir, "/ISIMIP2b/DerivedInputData/GCM/landonly"),
  pattern="bioclim_.*2050.*\\.csv.xz", full.names=T, recursive=T), list.files(
    paste0(filedir, "/ISIMIP2b/DerivedInputData/GCM/landonly"),
    pattern="bioclim_.*2080.*\\.csv.xz", full.names=T, recursive=T))
bioclim_fut <- lapply(bioclim_fut, function(x){
  data <- read.csv(x)
  data$year <- strsplit(basename(x), split="_")[[1]][4]
  data$model <- strsplit(basename(x), split="_")[[1]][2]
  data$scenario <- strsplit(basename(x), split="_")[[1]][3]
  return(data)
})
bioclim_fut <- do.call("rbind", bioclim_fut)

library(dplyr)
bioclim_fut <- bioclim_fut %>% 
  dplyr::select(c(x,y,model,scenario,year,bio4,bio5,bio12,bio15,bio18,bio19))
bioclim_fut <- tidyr::gather(bioclim_fut, var, value, -c(x,y,model,scenario,year))
bioclim_fut <- tidyr::spread(bioclim_fut, year, value)

# Calculate delta climate
bioclim_all <- left_join(bioclim_fut, bioclim_1995, by=c("x", "y", "var"))
delta_climate <- bioclim_all %>% 
  mutate_at(vars(`2050`:`2080`), funs(. - bioclim_all$`1995`)) %>% dplyr::select(-c(`1995`))
delta_climate <- tidyr::gather(delta_climate, year, value, -c(x,y,model,scenario,var))
delta_climate <- delta_climate %>% group_by(x,y,scenario,var, year) %>% 
  summarise(value=mean(value, na.rm=TRUE))

#Subset data for plotting
lapply(c("2050", "2080"), function(x){
  climate <- delta_climate[delta_climate$year == x,]
  climate <- climate %>% tidyr::unite(year, scenario, col="time_rcp")
  climate$time_rcp <- factor(climate$time_rcp, labels=c(paste0(x, " RCP2.6"), 
                                                        paste0(x, " RCP6.0")))
  data(outline, package="ggmap2")
  library(ggplot2)
  p1 <- ggplot() + 
    geom_raster(data=climate[climate$var == "bio4",], aes(x=x, y=y, fill=value)) + 
    facet_wrap(~ time_rcp, ncol=2) + 
    geom_polygon(data=outline, aes(x=long,y=lat, group=group), 
                 fill="transparent", colour="black") + 
    scale_fill_gradientn(name="Temperature \nseasonality", colours=rev(colorRampPalette(
      c("#00007F", "blue", "#007FFF", "cyan", "white", "yellow", 
        "#FF7F00", "red", "#7F0000"))(255)), values=scales::rescale(unique(c(seq(min(climate$value[climate$var == "bio4"]), 0, length=5), seq(0, max(climate$value[climate$var == "bio4"]), length=5)))), na.value="transparent") +
    theme_classic() + theme(axis.title = element_blank(), axis.line = element_blank(),
                            axis.ticks = element_blank(), axis.text = element_blank(),
                            panel.grid = element_blank(), 
                            strip.background= element_blank(),
                            strip.placement = "outside", 
                            strip.text = element_text(size=10, face="bold"),
                            rect = element_rect(fill = "transparent")) + 
    coord_quickmap(xlim=c(-180,180), ylim=c(-60,85), expand=FALSE)
  p2 <- ggplot() + 
    geom_raster(data=climate[climate$var == "bio12",], aes(x=x, y=y, fill=value)) + 
    facet_wrap(~ time_rcp, ncol=2) + 
    geom_polygon(data=outline, aes(x=long,y=lat, group=group), 
                 fill="transparent", colour="black") + 
    scale_fill_gradientn(name="Annual \nprecipitation", colours=colorRampPalette(
      c("#00007F", "blue", "#007FFF", "cyan", "white", "yellow", 
        "#FF7F00", "red", "#7F0000"))(255),
      values=scales::rescale(unique(c(seq(min(climate$value[climate$var == "bio12"]), 
                                          0, length=5), seq(0, max(climate$value[climate$var == "bio12"]), length=5)))), na.value="transparent") +
    theme_classic() + theme(axis.title = element_blank(),axis.line = element_blank(),
                            axis.ticks = element_blank(), axis.text = element_blank(),
                            panel.grid = element_blank(), 
                            strip.background= element_blank(),
                            strip.placement = "outside", 
                            strip.text = element_blank(),
                            rect = element_rect(fill = "transparent")) + 
    coord_quickmap(xlim=c(-180,180), ylim=c(-60,85), expand=FALSE)
  p3 <- ggplot() + 
    geom_raster(data=climate[climate$var == "bio19",], aes(x=x, y=y, fill=value)) + 
    facet_wrap(~ time_rcp, ncol=2) + 
    geom_polygon(data=outline, aes(x=long,y=lat, group=group), 
                 fill="transparent", colour="black") + 
    scale_fill_gradientn(name="Precipitation \nof coldest \nquarter", 
                         colours=colorRampPalette(
                           c("#00007F", "blue", "#007FFF", "cyan", "white", "yellow", 
                             "#FF7F00", "red", "#7F0000"))(255), values=scales::rescale(unique(c(seq(min(climate$value[climate$var == "bio19"]), 0, length=5), seq(0, max(climate$value[climate$var == "bio19"]), length=5)))), na.value="transparent") +
    theme_classic() + theme(axis.title = element_blank(),axis.line = element_blank(),
                            axis.ticks = element_blank(), axis.text = element_blank(),
                            panel.grid = element_blank(),
                            strip.background= element_blank(),
                            strip.placement = "outside", strip.text = element_blank(),
                            rect = element_rect(fill = "transparent")) + 
    coord_quickmap(xlim=c(-180,180), ylim=c(-60,85), expand=FALSE)
  
  # Turn plots into grob elements
  p <- lapply(list(p1,p2,p3), ggplotGrob)
  p <- rbind(p[[1]], p[[2]], p[[3]], size="first")
  for (i in which(p$layout$name == "guide-box")) {
    p$grobs[[i]] <- p$grobs[[i]]$grobs[[1]]
  }
  png(file=paste0("figures/top_clim_change_", x, ".png"), 
      width=10, height=6, unit="in", res=600)
  grid::grid.draw(p)
  dev.off()
})

lapply(c("2050", "2080"), function(time){
  climate <- delta_climate[delta_climate$year == time,]
  climate <- climate %>% tidyr::unite(year, scenario, col="time_rcp")
  climate$time_rcp <- factor(climate$time_rcp, labels=c(paste0(time, " RCP2.6"), 
                                                        paste0(time, " RCP6.0")))
  data(outline, package="ggmap2")
  library(ggplot2)
  p1 <- ggplot() + 
    geom_raster(data=climate[climate$var == "bio5",], aes(x=x, y=y, fill=value)) + 
    facet_wrap(~ time_rcp, ncol=2) + 
    geom_polygon(data=outline, aes(x=long,y=lat, group=group), 
                 fill="transparent", colour="black") + 
    scale_fill_gradientn(name="Maximum \ntemperature", colours=rev(colorRampPalette(
      c("#00007F", "blue", "#007FFF", "cyan", "white", "yellow", 
        "#FF7F00", "red", "#7F0000"))(255)),
      values=scales::rescale(unique(c(seq(min(climate$value[climate$var == "bio5"]), 0, length=5), seq(0, max(climate$value[climate$var == "bio5"]), length=5)))), na.value="transparent") +
    theme_classic() + theme(axis.title = element_blank(),axis.line = element_blank(),
                            axis.ticks = element_blank(), axis.text = element_blank(),
                            panel.grid = element_blank(), strip.background= element_blank(),
                            strip.placement = "outside", strip.text = element_text(size=10, face="bold"),
                            rect = element_rect(fill = "transparent")) + 
    coord_quickmap(xlim=c(-180,180), ylim=c(-60,85), expand=FALSE)
  p2 <- ggplot() + 
    geom_raster(data=climate[climate$var == "bio15",], aes(x=x, y=y, fill=value)) + 
    facet_wrap(~ time_rcp, ncol=2) + 
    geom_polygon(data=outline, aes(x=long,y=lat, group=group), 
                 fill="transparent", colour="black") + 
    scale_fill_gradientn(name="Precipitation \nseasonality", colours=colorRampPalette(
      c("#00007F", "blue", "#007FFF", "cyan", "white", "yellow", 
        "#FF7F00", "red", "#7F0000"))(255), values=scales::rescale(unique(c(seq(min(climate$value[climate$var == "bio15"]), 0, length=5), seq(0, max(climate$value[climate$var == "bio15"]), length=5)))), na.value="transparent") +
    theme_classic() + theme(axis.title = element_blank(),axis.line = element_blank(),
                            axis.ticks = element_blank(), axis.text = element_blank(),
                            panel.grid = element_blank(), strip.background= element_blank(),
                            strip.placement = "outside", strip.text = element_blank(),
                            rect = element_rect(fill = "transparent")) + 
    coord_quickmap(xlim=c(-180,180), ylim=c(-60,85), expand=FALSE)
  p3 <- ggplot() + 
    geom_raster(data=climate[climate$var == "bio18",], aes(x=x, y=y, fill=value)) + 
    facet_wrap(~ time_rcp, ncol=2) + 
    geom_polygon(data=outline, aes(x=long,y=lat, group=group), 
                 fill="transparent", colour="black") + 
    scale_fill_gradientn(name="Precipitation \nof warmest \nquarter", 
                         colours=colorRampPalette(
                           c("#00007F", "blue", "#007FFF", "cyan", "white", "yellow", 
                             "#FF7F00", "red", "#7F0000"))(255), values=scales::rescale(unique(c(seq(min(climate$value[climate$var == "bio18"]), 0, length=5), seq(0, max(climate$value[climate$var == "bio18"]), length=5)))), na.value="transparent") +
    theme_classic() + theme(axis.title = element_blank(),axis.line = element_blank(),
                            axis.ticks = element_blank(), axis.text = element_blank(),
                            panel.grid = element_blank(), strip.background= element_blank(),
                            strip.placement = "outside", strip.text = element_blank(),
                            rect = element_rect(fill = "transparent")) + 
    coord_quickmap(xlim=c(-180,180), ylim=c(-60,85), expand=FALSE)
  
  # Turn plots into grob elements
  p <- lapply(list(p1,p2,p3), ggplotGrob)
  p <- rbind(p[[1]], p[[2]], p[[3]], size="first")
  for (i in which(p$layout$name == "guide-box")) {
    p$grobs[[i]] <- p$grobs[[i]]$grobs[[1]]
  }
  png(file=paste0("figures/low_clim_change_", time, ".png"), 
      width=10, height=6, unit="in", res=600)
  grid::grid.draw(p)
  dev.off()
})
```

### Create line graph of the different temperature scenarios

```{r, boxplot_bio}
#' Create maps and plots of climate data

filedir <- "/home/mabi/Documents/Wissenschaft/Data"
filedir <- "E:/Data"

# Get climate files (csv with xy and BioClimVar)
clim_data <- list.files(paste0(filedir, "/ISIMIP2b/DerivedInputData/GCM/landonly"), pattern="bioclim",
                        recursive=TRUE, full.names=TRUE)

# Retrieve information on Model, RCP and timeframe
models <- lapply(clim_data, function(x) strsplit(basename(x), split="_")[[1]][2])
rcps <- lapply(clim_data, function(x) strsplit(basename(x), split="_")[[1]][3])
times <- lapply(clim_data, function(x) strsplit(basename(x), split="[.;_]")[[1]][4])

# Read in future climate files
clim_data <- lapply(clim_data, function(x) readr::read_csv(x))

# Add column for model type, rcp and timeframe to each list
clim_data<- Map(cbind, clim_data, model = models, rcp = rcps, time = times)

# Turn climate data into dataframe
clim_data <- do.call("rbind", clim_data)

# Adjust EWEMBI data
str(clim_data)
clim_data$time <- as.character(clim_data$time)
clim_data$time[clim_data$model == "EWEMBI"] <- 1995
clim_data$time <- as.numeric(clim_data$time)
clim_data$rcp <- as.character(clim_data$rcp)
clim_data$rcp[clim_data$model == "EWEMBI"] <- "rcp26"
clim_ewembi <- clim_data[clim_data$model == "EWEMBI",]
clim_ewembi$rcp <- "rcp60"
clim_data <- rbind(clim_data, clim_ewembi); rm(clim_ewembi)

# Only select bio4, bio5, bio12, bio15, bio18 bio19
library(dplyr); library(tidyr)
clim_data <- clim_data %>% select(x, y, bio4, bio5, bio12, bio15, bio18, 
                                  bio19, model, rcp, time) %>%
  gather(key=var, value=value, -c(x, y, model, rcp, time))
clim_data$var <- factor(clim_data$var, 
                        levels=c("bio4", "bio5", "bio12", "bio15", "bio18", "bio19"),
                        labels=c("Bio 4", "Bio 5", "Bio 12", "Bio 15", "Bio 18", "Bio 19"))

library(rISIMIP)
data("landseamask_generic", package="rISIMIP")
area <- raster::area(landseamask_generic)
area <- as.data.frame(rasterToPoints(area))
colnames(area) <- c("x", "y", "area")

# Add area of each cell to clim_data
isimip_area <- isimip_area
clim_data <- left_join(clim_data, isimip_area, by=c("x", "y"))

# Create line graph
clim_data <- left_join(clim_data, area, by=c("x", "y"))
clim_mean <- clim_data %>% filter(time %in% c(1995, 2020, 2050, 2080)) %>% 
  group_by(model, rcp, time, var) %>% summarise(mean=weighted.mean(value, w=area, na.rm=TRUE))
ggplot(data=clim_mean, 
       aes(x=time, y=mean, 
           colour=factor(model), linetype=factor(rcp))) + geom_point() + 
  geom_line() +
  labs(x="", y="") +
  facet_wrap(~ var, scales="free_y", ncol=2) + 
  scale_x_continuous(breaks=c(1995,2020, 2050,2080)) +
  scale_colour_discrete(name="Model") + 
  scale_linetype_discrete(name="Scenario") + 
  theme_bw() + theme(strip.background= element_blank(),
                     legend.background = element_rect(fill = NA),
                     panel.spacing.x=unit(0.25, "lines"),
                     panel.spacing.y=unit(0.25, "lines"))
ggsave("figures/model_variables_1995_2080.pdf", width=9, height=6, dpi=600)

# Load GitHub package
library(ggmap2)

#Plot Map of current climate data
clim_data_wide <- 
  createMap(data=clim_data_wide, name="Bio4")

# Plot Map of future climate data
climSub <- clim[c(1,2,6,22,23,24)]
climSub <- climSub[climSub$rcp == "rcp26",]
colnames(climSub) <- c("x", "y", "value", "var", "rcp", "var2")
createMap(data=climSub, split=FALSE, facet_grid = TRUE, long=TRUE)
```
