---
title: "Global mean - ISIMIP2b"
author: "RS-eco"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Global mean - ISIMIP2b}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, warning=F, comment=NA, message=F, eval=F)
```

## Data Setup

Load packages

```{r}
library(rISIMIP)
library(ggplot2)
library(dplyr)
```

First, specify the file path, where the ISIMIP data is located.

```{r global_options}
# Specify path of file directory
filedir <- "/media/matt/Data/Documents/Wissenschaft/Data/"
```

## Mean global annual mean ISIMIP2b temperature

This data can be downloaded from the ISIMIP Server and is included in this package.

```{r, eval=FALSE}
data(yearmean_tas, package="rISIMIP")
data(runmean31_tas, package="rISIMIP")
```

## Mean global annual mean EWEMBI temperature

Load EWEMBI mean temperature files and calculate annual global mean

```{r, tasmean_EWEMBI, fig.width=9, fig.height=6, dpi=600}
# Load temperature files
tas_ewembi_files <- list.files(
  paste0(filedir, "/ISIMIP2b/EWEMBI"),
  pattern="^tas_", full.names=TRUE, recursive=T)
outfiles <- sub("tas_", "yearmean_tas_", sub("EWEMBI", "DerivedInputData/globalmeans/EWEMBI", tas_ewembi_files))

landseamask <- paste0(filedir, "ISIMIP2b/InputData/landseamask/ISIMIP2b_landseamask_generic.nc4")
mapply(FUN=function(x,y) {
  if(!file.exists(y)){system(paste0('cdo fldmean -yearmean -ifthen ', landseamask, " ", x, ' ', y))}}, x=tas_ewembi_files, y=outfiles); rm(tas_ewembi_files, landseamask)  

# Read outfiles into one data.frame
tas <- lapply(outfiles, function(x){
  nc <- ncdf4::nc_open(x)
  tas <- ncdf4::ncvar_get(nc)
  tas <- as.data.frame(tas)
  timeref <- as.Date(strsplit(nc$dim[[1]]$units, " ")[[1]][3]) 
  if(ncdf4::ncvar_get(nc, nc$dim$time)[1] == 0){
    tas$year <- timeref + ncdf4::ncvar_get(nc, nc$dim$time)
  } else if (ncdf4::ncvar_get(nc, nc$dim$time)[1] != 0){
    tas$year <- timeref + ncdf4::ncvar_get(nc, nc$dim$time) - 1
  }
  ncdf4::nc_close(nc)
  return(tas)
})
tas <- dplyr::bind_rows(tas); rm(outfiles)

# Create plot
tas$year <- as.Date(tas$year)
library(ggplot2)
ggplot(data=tas, aes(x=year, y=tas-273.15)) + geom_line() + 
  labs(x="Year", y="Mean global annual mean temperature (°C)") + 
  scale_x_date() + theme_bw() + theme(strip.background= element_blank(),
                     strip.placement= "outside",
                     legend.position = c(0.9,0.905),
                     legend.title = element_blank(),
                     legend.background = element_rect(fill = NA),
                     panel.spacing.x=unit(0.25, "lines"),
                     panel.spacing.y=unit(0.25, "lines"))
```

## Mean global annual mean temperature

```{r}

# Compare mean 31-yr average temperature for 2050, 2080 among RCPs
library(dplyr)
runsub <- runmean31_tas %>% filter(scenario %in% c("rcp26", "rcp60")) %>%
  filter(year %in% c("2050", "2080")) %>% group_by(year, scenario, model)
colnames(runsub)[2] <- "tas"
runsub$tas <- runsub$tas - 286.7891

# Calculate 31yr running mean from EWEMBI data
tas$year <- lubridate::year(tas$year)
runmean31_tas <- zoo::rollapply(zoo::as.zoo(tas[,c("tas", "year")]), 31, mean, na.rm=TRUE, partial=TRUE)
runmean31_tas <- as.data.frame(runmean31_tas)
runmean31_tas[runmean31_tas$year == 1995,]
```

## Mean global annual maximum temperature

```{r}
# List tasmax files
tasmax_files <- list.files(
  paste0(filedir, "/ISIMIP2b/InputData/landonly"),
  pattern="^tasmax_.*.nc4", full.names=TRUE, recursive=T)
# Create outfile names
outfiles <- sub("tasmax_day", "yearmax_tas", sub("InputData/landonly", "DerivedInputData/globalmeans/", tasmax_files))

# Load landsea mask
landseamask <- paste0(filedir, "ISIMIP2b/InputData/landseamask/ISIMIP2b_landseamask_generic.nc4")

# Calculate global mean annual maximum temperature
mapply(FUN=function(x,y) {if(!file.exists(y)){system(paste0('cdo fldmean -yearmax -ifthen ', landseamask, " ", x, ' ', y))}}, x=tasmax_files, y=outfiles)
# Without landsea subset
mapply(FUN=function(x,y) {if(!file.exists(y)){system(paste0('cdo fldmean -yearmax ', x, ' ', y))}}, x=tasmax_files, y=outfiles)

# Save outfiles in csv file
yearmax_tas <- lapply(outfiles, function(x){
  nc <- ncdf4::nc_open(x)
  tasmax <- ncdf4::ncvar_get(nc)
  tasmax <- as.data.frame(tasmax)
  timeref <- as.Date(strsplit(nc$dim[[1]]$units, " ")[[1]][3]) 
  if(ncdf4::ncvar_get(nc, nc$dim$time)[1] == 0){
    tasmax$year <- timeref + ncdf4::ncvar_get(nc, nc$dim$time)
  } else if (ncdf4::ncvar_get(nc, nc$dim$time)[1] != 0){
    tasmax$year <- timeref + ncdf4::ncvar_get(nc, nc$dim$time) - 1
  }
  ncdf4::nc_close(nc)
  tasmax$model <- strsplit(x, split="_")[[1]][3]
  tasmax$scenario <- strsplit(x, split="_")[[1]][4]
  return(tasmax)
})
yearmax_tas <- do.call(rbind, tasmax)
```

## Mean global annual total precipitation

```{r}
pr_files <- list.files(
  paste0(filedir, "/ISIMIP2b/InputData/landonly"),
  pattern="^pr_.*.nc4", full.names=TRUE, recursive=T)
outfiles <- sub("pr_day", "yearsum_pr", sub("InputData/landonly", "DerivedInputData/globalmeans", pr_files))

landseamask <- paste0(filedir, "ISIMIP2b/InputData/landseamask/ISIMIP2b_landseamask_generic.nc4")

mapply(FUN=function(x,y) {if(!file.exists(y)){system(paste0('cdo fldmean -yearsum -ifthen ', landseamask, " ", x, ' ', y))}}, x=pr_files, y=outfiles)
# Without landsea subset
mapply(FUN=function(x,y) {if(!file.exists(y)){system(paste0('cdo fldmean -yearsum ', x, ' ', y))}}, x=pr_files, y=outfiles)

# Save outfiles in csv file
yearsum_pr <- lapply(outfiles, function(x){
  nc <- ncdf4::nc_open(x)
  pr <- ncdf4::ncvar_get(nc)
  pr <- as.data.frame(pr)
  timeref <- as.Date(strsplit(nc$dim[[1]]$units, " ")[[1]][3]) 
  if(ncdf4::ncvar_get(nc, nc$dim$time)[1] == 0){
    pr$year <- timeref + ncdf4::ncvar_get(nc, nc$dim$time)
  } else if (ncdf4::ncvar_get(nc, nc$dim$time)[1] != 0){
    pr$year <- timeref + ncdf4::ncvar_get(nc, nc$dim$time) - 1
  }
  ncdf4::nc_close(nc)
  pr$model <- strsplit(x, split="_")[[1]][3]
  pr$scenario <- strsplit(x, split="_")[[1]][4]
  return(pr)
})
yearsum_pr <- do.call(rbind, yearsum_pr)
```

## Calculate 31-year mean

```{r runmean31}
# Calculate 31yr running mean
runmean31_tasmax <- lapply(unique(yearmax_tas$model), function(x){
 data_agg <- lapply(unique(yearmax_tas$scenario), function(y){
   data <- yearmax_tas[yearmax_tas$model == x,]
   data <- data[data$scenario == y,]
   data$year <- lubridate::year(data$year)
   data_mean <- zoo::rollapply(zoo::as.zoo(data[,c("tasmax", "year")]), 31, mean, na.rm=TRUE, partial=TRUE)
   data_mean <- as.data.frame(data_mean)
   data_mean$scenario <- y
   return(data_mean)
 })
 data_agg <- do.call("rbind", data_agg)
 data_agg$model <- x
 return(data_agg)
})
runmean31_tasmax <- do.call("rbind", runmean31_tasmax)

# Calculate 31yr running mean
runmean31_prsum <- lapply(unique(yearsum_pr$model), function(x){
 data_agg <- lapply(unique(yearsum_pr$scenario), function(y){
   data <- yearsum_pr[yearsum_pr$model == x,]
   data <- data[data$scenario == y,]
   data$year <- lubridate::year(data$year)
   data_mean <- zoo::rollapply(zoo::as.zoo(data[,c("pr", "year")]), 31, mean, na.rm=TRUE, partial=TRUE)
   data_mean <- as.data.frame(data_mean)
   data_mean$scenario <- y
   return(data_mean)
 })
 data_agg <- do.call("rbind", data_agg)
 data_agg$model <- x
 return(data_agg)
})
runmean31_prsum <- do.call("rbind", runmean31_prsum)
```

### Plots

```{r tasmax_change, fig.width=9, fig.height=6, dpi=600}
# Create simple plot 
yearmax_tas$year <- lubridate::year(yearmax_tas$year)
ggplot(data=yearmax_tas, aes(x=year, y=tasmax-273.15, colour=scenario)) + geom_line() + 
  facet_wrap(~ model, scales="free_y", ncol=1) + 
  #facet_grid(var~model, scales="free_y", switch="y") + 
  labs(x="Year", y="Mean global annual maximum temperature (°C)") + 
  scale_x_continuous(limits=c(1650, 2310), expand=c(0,0),
                     breaks=c(1700, 1800, 1900, 2000, 2100, 2200, 2300)) + 
  #scale_colour_manual(values=c("black", "darkgrey", "blue", "yellow")) + 
  theme_bw() + theme(strip.background= element_blank(),
                     strip.placement= "outside",
                     legend.position = c(0.9,0.905),
                     legend.title = element_blank(),
                     legend.background = element_rect(fill = NA),
                     panel.spacing.x=unit(0.25, "lines"),
                     panel.spacing.y=unit(0.25, "lines"))
```

```{r prsum_change, fig.width=9, fig.height=6, dpi=600}
# Create simple plot 
yearsum_pr$year <- lubridate::year(yearsum_pr$year)
ggplot(data=yearsum_pr, aes(x=year, y=pr*86400, colour=scenario)) + geom_line() + 
  facet_wrap(~ model, scales="free_y", ncol=1) +
  labs(x="Year", y="Mean global annual total precipitation (mm)") + 
  #scale_x_date() + 
  scale_x_continuous(limits=c(1650, 2310), expand=c(0,0), 
                     breaks=c(1700, 1800, 1900, 2000, 2100, 2200, 2300)) +
  #scale_colour_manual(values=c("black", "darkgrey", "blue", "yellow")) + 
  theme_bw() + theme(strip.background= element_blank(),
                     strip.placement= "outside",
                     legend.position = c(0.9,0.905),
                     legend.title = element_blank(),
                     legend.background = element_rect(fill = NA),
                     panel.spacing.x=unit(0.25, "lines"),
                     panel.spacing.y=unit(0.25, "lines"))
```

```{r pr_tasmax_ipsl-cm5a-lr, width=9, height=6, dpi=600}
library(dplyr)
yearsum_ipsl <- yearsum_pr %>% filter(model == "IPSL-CM5A-LR")
yearsum_ipsl$var <- "pr"
colnames(yearsum_ipsl)[1] <- "value"
yearsum_ipsl$value <- yearsum_ipsl$value*86400
yearmax_ipsl <- yearmax_tas %>% filter(model == "IPSL-CM5A-LR")
yearmax_ipsl$var <- "tasmax"
colnames(yearmax_ipsl)[1] <- "value"
yearmax_ipsl$value <- yearmax_ipsl$value-273.15
yearvar_ipsl <- bind_rows(yearsum_ipsl, yearmax_ipsl)
ggplot(data=yearvar_ipsl, aes(x=year, y=value, colour=scenario)) + 
  geom_line() + facet_wrap(~ var, scales="free_y", ncol=1,
                           strip.position="left", labeller= as_labeller(c(pr="Mean global annual total precipitation (mm)", tasmax="Mean global annual maximum temperature (°C)"))) +
  labs(x="Year", y="") + 
  scale_x_date() + 
  #scale_colour_manual(values=c("black", "darkgrey", "blue", "yellow")) + 
  theme_bw() + theme(strip.background= element_blank(),
                     strip.placement= "outside",
                     legend.position = c(0.06,0.9),
                     legend.title = element_blank(),
                     legend.background = element_rect(fill = NA),
                     panel.spacing.x=unit(0.25, "lines"),
                     panel.spacing.y=unit(0.25, "lines"))
```
