---
title: "Introduction to rISIMIP"
author: "RS-eco"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Introduction to rISIMIP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=14, fig.height=8, warning=FALSE, comment=NA, message=FALSE, eval=F)
```

## Data Set-Up

Load rISIMIP package

```{r}
library(rISIMIP)
```

First, we specify the file path, where the ISIMIP data is located.

```{r global_options}
# Specify path of file directory
filedir <- "/mnt/sda1/Documents/Wissenschaft/Data" 
# choose.dir() # Only works on Windows
```

## Data Pre-processing

### Climate data

#### Read climate files

With `readISIMIP` you can read multiple ISIMIP datafiles into one raster stack.

```{r read_climate, eval=FALSE}
# GCM, CM5A-R, historical and RCP2.6, 1970 - 2000 (Current conditions)
tas_1991_2000 <- readISIMIP(path=filedir, var="tas", model="GFDL-ESM2M", 
                      startyear=1991, endyear=2000)
```

However, this is not useful if you are interested in long time periods, as one datafile is about 7 GB in size.

#### List climate files

For daily data please use the `summariseNC` function in the `processNC` package, if you require summarised time periods.

The function listISIMIP only lists all climate files for the desired time period, model and variable. The files can then be put into the `summariseNC` function for processing the required NetCDF files.

```{r list_climate}
# List hurs files for EWEMBI from 1970 - 1999
hurs_1995 <- listISIMIP(path=filedir, var="hurs", model="EWEMBI",
                       startyear=1980, endyear=2009)
hurs_1995
```

### Summarise climate data

First, we need the `summariseNC` function from the `processNC` package.
Thus, we have to install the processNC package from Github.

```{r install_processNC, eval=FALSE}
# Install processNC package
devtools::install_github("RS-eco/processNC")
```

Load processNC package

```{r}
library(processNC)
```

Now we can summarise our data using the `summariseNC` function.

```{r summarise_climate, eval=FALSE}
# Create timeframe summaries
mean_hurs_1995 <- summariseNC(files=hurs_1995, filename1="monthly_hurs_1995.grd", 
                             format="raster", overwrite=TRUE)
```

We now list and summarise the climate data for the required time steps (1970-1999, 2006-2035, 2036-2065, 2066-2095, 2086-2115, 2136-2165) for each variable and each model using the global data.

## Data Processing

### Transform climate data

#### Change units

The climate data comes in non-standard units. Temperature is in Kelvin and needs to be converted to ° Celsius, while precipitation was originally in kg m-2 s-1 and needs to be converted to kg m-2 day-1, which equals mm per day.

```{r}
# Turn climate data into right units (°C and mm)

# List all temperature (tas, tasmin, tasmax) files
tas_files <- list.files(paste0(filedir, "/ISIMIP2b/ProcessedData/global/ewembi/"), 
                  pattern="monthly_tas_.*\\.grd", full.names=TRUE)
tmin_files <- list.files(paste0(filedir, "/ISIMIP2b/ProcessedData/global/ewembi/"), 
                  pattern="monthly_tasmin_.*\\.grd", full.names=TRUE)
tmax_files <- list.files(paste0(filedir, "/ISIMIP2b/ProcessedData/global/ewembi/"), 
                   pattern="monthly_tasmax_.*\\.grd", full.names=TRUE)

#Convert temperature from Kelvin to ° Celsius
tas <- lapply(tas_files, FUN=function(x){
  x <- raster::stack(x)
  raster::calc(x, fun=function(x){x-273.15})
})
tmin <- lapply(tmin_files, FUN=function(x){
  x <- raster::stack(x)
  raster::calc(x, fun=function(x){x-273.15})
})
tmax <- lapply(tmax_files, FUN=function(x){
  x <- raster::stack(x)
  raster::calc(x, fun=function(x){x-273.15})
})

# List precipitation files
prec_files <-  list.files(paste0(filedir, "/ISIMIP2b/ProcessedData/global/ewembi/"), 
                    pattern="monthly_pr_.*\\.grd", full.names=TRUE)

# Convert precipitation from kg m-2 s-1 to kg m-2 day-1
prec <- lapply(prec_files, FUN=function(x){
  x <- raster::stack(x)
  raster::calc(x, fun=function(x){x*86400})
})
```

### Calculate Bioclim variables

From the resulting layers, we can now calculate the Bioclimatic variables, using the biovars function in the dismo package.

```{r bioclim, eval=FALSE}
library(dismo)

# Create list with bioclim names
bioclim_names <- gsub(x = basename(prec_files), pattern = "\\monthly_pr", replacement = "bioclim")

# Calculate bioclim variables for all models and time frames and save to file
bioclim <- mapply(FUN=function(x,y,z,name){
  bio <- dismo::biovars(tmin=x, tmax=y, prec=z)
  writeRaster(bio, filename=paste0(filedir, "/ISIMIP2b/ProcessedData/global/bioclim/", name), format="raster", overwrite=FALSE)
  return(bio)
  }, x=tmin, y=tmax, z=prec, name=bioclim_names)
```

Now, we could also list previously calculated bioclim files

```{r}
bioclim_files <- list.files(paste0(filedir, "/ISIMIP2b/ProcessedData/global/bioclim/"), 
                  pattern="bioclim_.*\\.grd", full.names=TRUE)
bioclim <- lapply(bioclim_files, stack)
```

## Data visualisation

### ggmap2 package

For plotting our data with the `createMap` function, we need to install the `ggmap2` package.

```{r, eval=FALSE}
# Install ggmap2 package from Github
devtools::install_github("RS-eco/ggmap2")
```

Load ggmap2 package

```{r}
library(ggmap2)
```

# Add plot example!!!

### Plot Bioclim data

```{r, bioclim_2050, eval=FALSE}
# Read Bioclim data files
bioclim_2050 <- lapply(list.files(paste0(filedir, "ISIMIP2b/ProcessedData/bioclim"), 
                                pattern="rcp_2050.grd", full.names=TRUE), stack)

# Create Map of bioclim data
bio04_2050 <- stack(lapply(bioclim_2050, function(x) x[[4]]))
createMap(bioclim[[20]][[4]], name="Bio04", split=FALSE, ncol=2, width=8, height=12, units="in", filename=NA, dpi=100)
```

## Landuse data

### Get landuse data

```{r landuse, eval=FALSE}
# Landuse histsoc
crops5_1985 <- readISIMIP(path=filedir, type="landuse", scenario="histsoc", 
                          var="5crops", startyear=1970, endyear=1999)
crops15_1985 <- readISIMIP(path=filedir, type="landuse", scenario="histsoc", 
                           var="15crops", startyear=1970, endyear=1999)
totals_1985 <- readISIMIP(path=filedir, type="landuse", scenario="histsoc", 
                          var="totals", startyear=1970, endyear=1999)
urban_1985 <- stack(readISIMIP(path=filedir, type="landuse", scenario="histsoc", 
                         var="urbanareas", startyear=1970, endyear=1999))
```

### Summarise Landuse data

```{r summarise_lu, eval=FALSE}
# Reference data
crops5_1985 <- stack(lapply(crops5_1985, FUN=function(x) calc(x, mean)))
crops15_1985 <- stack(lapply(crops15_1985, FUN=function(x) calc(x, mean)))
totals_1985 <- stack(lapply(totals_1985, FUN=function(x) calc(x, mean)))
urban_1985 <- calc(urban_1985, mean)
```

### Plot land use data

```{r, eval=FALSE}
createMap(totals_1985, name="% Cover", subnames=names(crops5_1985), split=FALSE, ncol=4, width=12, height=8, units="in", dpi=100)
```

### Save Landuse data to file

```{r, eval=FALSE}
lu_files <- list(crops5_1985, crops15_1985, totals_1985, urban_1985)
filenames <- c("crops5_histsoc_1985.tif", "crops15_histsoc_1985.tif", "totals_histsoc_1985.tif", "urbanareas_histsoc_1985.tif")
mapply(FUN=function(x,y) writeRaster(x=x, filename=paste0("extdata/", y), format="GTiff", overwrite=T), x=lu_files, y=filenames)
```

## Present scenario

```{r, eval=FALSE}
# Landuse 2005soc - future time periods
totals_2005soc <- stack(readISIMIP(path=filedir, type="landuse", scenario="2005soc", var="totals", startyear=2006, endyear=2007))[[c(1,3,5,7)]]
urban_2005soc <- stack(readISIMIP(path=filedir, type="landuse", scenario="2005soc", 
                         var="urbanareas", startyear=2006, endyear=2007))[[1]]

lu_files <- list(totals_2005soc, urban_2005soc)
filenames <- c("totals_2005soc.tif", "urbanareas_2005soc.tif")
mapply(FUN=function(x,y) writeRaster(x=x, filename=paste0("extdata/", y), format="GTiff", overwrite=T), x=lu_files, y=filenames)

lu_files <- list(totals_2005soc, urban_2005soc)
filenames <- c("totals_2005soc.csv", "urbanareas_2005soc.csv")
colnames <- list(c("x", "y", "cropland_total", "pastures", "cropland_irrigated", "cropland_rainfed"), c("x", "y", "urbanareas"))
mapply(FUN=function(x,y,z){
  data <- as.data.frame(rasterToPoints(x))
  colnames(data) <- z
  readr::write_csv(x=data, path=paste0("extdata/", y))}, 
  x=lu_files, y=filenames,z=colnames)
```

## Future scenarios

For the different future timeframes, we now do this all in one.

```{r lu_future, eval=FALSE, echo=FALSE}
# Time frames rcp26
timeframes <- c("2020","2050","2080","2100","2150")
startyears <- c(2006,2036,2066,2086,2136)
endyears <- c(2035,2065,2095,2115,2165)

# Only totals land use data is available
totals_rcp26_all <- mapply(FUN=function(x,y){
  data <- readISIMIP(path=filedir, type="landuse", scenario="rcp26",
                     var="totals", startyear=x, endyear=y)
  data <- stack(lapply(data, FUN=function(x) calc(x, mean)))
  return(data)
}, startyears, endyears)
names(totals_rcp26_all) <- timeframes

# Urban data
urban_rcp26_all <- mapply(FUN=function(x,y){
  data <- stack(readISIMIP(path=filedir, type="landuse", scenario="rcp26",
                         var="urbanareas", startyear=x, endyear=y))
  data <- calc(data, mean)
  return(data)
}, startyears, endyears)
names(urban_rcp26_all) <- timeframes

# Time frames rcp60
timeframes <- c("2020","2050","2080")
startyears <- c(2006,2036,2066)
endyears <- c(2035,2065,2095)

# Only totals land use data is available
totals_rcp60_all <- mapply(FUN=function(x,y){
  data <- readISIMIP(path=filedir, type="landuse", scenario="rcp60soc",
                     var="totals", startyear=x, endyear=y)
  data <- stack(lapply(data, FUN=function(x) calc(x, mean)))
  return(data)
}, startyears, endyears)
names(totals_rcp60_all) <- timeframes

# Urban data
urban_rcp60_all <- mapply(FUN=function(x,y){
  data <- stack(readISIMIP(path=filedir, type="landuse", scenario="rcp60soc", var="urbanareas", startyear=x, endyear=y))
  data <- calc(data, mean)
  return(data)
}, startyears, endyears)
names(urban_rcp60_all) <- timeframes
```

### Plot data

```{r, eval=FALSE, echo=FALSE}
# Create boxplot of the different scenarios
library(ggplot2)
ggplot() + geom_boxplot(data=landuse_data, aes(x=year, y=value, fill=var,linetype=scenario))
ggsave("landuse_scenarios.png", dpi=300, width=12, height=6)
```

## Population data

We read and summarise global population data for the different time periods.
Histsoc population data is different for every year, but 2005soc keeps the population constant throughout time.

```{r pop_data, eval=FALSE}
# Timeframes (Horizon 2050, 2080)
startyears <- c(2036,2066)
endyears <- c(2065,2095)

population_ref <- calc(readISIMIP(path=filedir, type="population", scenario="histsoc",
                             startyear=1970, endyear=1999), mean)
population_2005 <- stack(readISIMIP(path=filedir, type="population", scenario="2005soc", startyear=2010, endyear=2020))[[1]]
# 2005soc layers are the same for every year, so no point in summarising them.

# ssp2soc only goes until 2100
population_ssp2soc <- stack(mapply(FUN=function(x,y){calc(readISIMIP(path=filedir, type="population", scenario="ssp2soc", startyear=x, endyear=y), mean)}, startyears, endyears))

# Merge all population data
population_data <- stack(population_ref, population_2005, population_ssp2soc)
population_data <- setZ(population_data, c("1985", "2005", "2050","2080"), name="time")
```

### Visualise population data

```{r, eval=FALSE}
createMap(population_data, name="Population", split=TRUE, subnames=c("1985", "2005", "2050", "2080"),
          filename="figures/population_all.tiff", ncol=1, width=8, height=12, 
          units="in", dpi=100)
```
