---
title: "ISIMIP rasterized WDPA Data"
author: "RS-eco"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ISIMIP rasterized WDPA Data}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width=8, fig.height=6, warning=FALSE, comment=NA, message=FALSE, eval=F)
```

**Note: Here we use the wdpa data that is included in the rISIMIP package. To see how we created these data files, please have a look at the wdpa.R in the data-raw folder in the rISIMIP Github repository.

## Plot WDPA Data

Map of percentage of current PAs (2018)

```{r}
library(rISIMIP)
data("protectedareas_annual_1819_2018_landonly", package="rISIMIP")
data(outline, package="ggmap2") # The ggmap2 package is available on Github: http://github.com/RS-eco/ggmap2
outline <- sf::st_as_sf(outline)
library(ggplot2); library(dplyr)
protectedareas_annual_1819_2018_landonly %>%
  ggplot() + geom_tile(aes(x=x,y=y,fill=`2018`*100)) + 
  geom_sf(data=outline, fill=NA, colour="grey50") + 
  scale_fill_gradientn(name="", colours=colorRampPalette(c("white", "#00007F", "blue", "#007FFF", "cyan", 
                              "#7FFF7F", "yellow", "#FF7F00", "red", 
                              "#7F0000"))(255)) + 
  coord_sf(xlim=c(-160,160), ylim=c(-55,85)) + labs(x="", y="") + theme_bw()
```

Map of percentage PAs by 2018 for each IUCN Category

```{r}
data("protectedareas_iucn_cat_2018_landonly", package="rISIMIP")
protectedareas_iucn_cat_2018_landonly %>% tidyr::gather(iucn_cat, perc, -c(x,y)) %>% tidyr::drop_na() %>%
  ggplot() + geom_tile(aes(x=x,y=y,fill=perc*100)) + facet_wrap(.~iucn_cat) + 
  geom_sf(data=outline, fill=NA, colour="grey50") + 
  scale_fill_gradientn(name="", colours=colorRampPalette(c("white", "#00007F", "blue", "#007FFF", "cyan", 
                              "#7FFF7F", "yellow", "#FF7F00", "red", 
                              "#7F0000"))(255)) + 
  coord_sf(xlim=c(-160,160), ylim=c(-55,85)) + labs(x="", y="") + theme_bw()
```

Plot bar chart of total area of PAs per IUCN category

```{r}
data("landseamask_generic", package="rISIMIP")
area_df <- as.data.frame(raster::rasterToPoints(raster::area(landseamask_generic)))
protectedareas_iucn_cat_2018_landonly %>%  tidyr::gather(iucn_cat, perc, -c(x,y)) %>% tidyr::drop_na() %>%
  left_join(area_df) %>% group_by(iucn_cat) %>% summarise(total=sum(perc*100*layer/1000000,na.rm=T)) %>% 
  ggplot() + geom_bar(aes(x=iucn_cat,y=total, fill=iucn_cat), stat="identity") + theme_bw() + 
  scale_fill_discrete(name="") + labs(x="IUCN Category", y="Area (km2)") + 
  theme(legend.position = "none") + scale_y_continuous(limits=c(0,400), expand=c(0,0))
```

Plot PAs over time

```{r}
data("protectedareas_annual_1819_2018_landonly", package="rISIMIP")
protectedareas_annual_1819_2018_landonly %>% tidyr::gather(year, perc, -c(x,y)) %>% tidyr::drop_na() %>%
  filter(year %in% c(1975, 1980, 1985, 1990, 1995, 2000, 2005, 2010, 2015)) %>% 
  ggplot() + geom_tile(aes(x=x,y=y,fill=perc*100)) + facet_wrap(.~year) + 
  geom_sf(data=outline, fill=NA, colour="grey50") + 
  scale_fill_gradientn(name="", colours=colorRampPalette(c("white", "#00007F", "blue", "#007FFF", "cyan", 
                              "#7FFF7F", "yellow", "#FF7F00", "red", 
                              "#7F0000"))(255)) + 
  coord_sf(xlim=c(-160,160), ylim=c(-55,85)) + labs(x="", y="") + theme_bw()
```

Plot line graph of total area of PAs over time

```{r}
protectedareas_annual_1819_2018_landonly %>% tidyr::gather(year, perc, -c(x,y)) %>% 
  left_join(area_df) %>% group_by(year) %>% summarise(total=sum(perc*100*layer/1000000,na.rm=T)) %>% 
  ggplot() + geom_line(aes(x=as.numeric(year),y=total)) + theme_bw() + 
  labs(x="Year", y="Area (km2)") + scale_y_continuous(limits=c(-10,1200), expand=c(0,0))
```
